```{r}
library("repr")
options(warn=-1)# repr.plot.width=14, repr.plot.height=8) 
```

TO DO

- introduction about methods, few formulas
- algorithm presentation
- generate data, comparison with "normal" histogram
- application on spectrum
- different prior studies

```{r}
bayesian_blocks <- function(xs, prior=7.61, nn_vec=rep(1,length(xs))){
    
    
    xs <- sort(xs)
    begin <- xs[1]
    end <- xs[length(xs)]
    
    xs <- xs[!(nn_vec==0)]
    nn_vec <- nn_vec[!(nn_vec==0)]
    
    N <- length(xs)
    start <- xs[1]
    stop <- xs[N]
  
    blocks <- c(start, 0.5*(xs[2:N]+xs[1:N-1]), stop);
    block_length <- stop - blocks
    # ---------------------------------------------
    # Start with first data cell; add one cell at
    # each iteration
    # ---------------------------------------------

    best <- NULL 
    last <- NULL
    supp <- NULL
    
    for (R in 1:N){
    # Compute fit_vec : fitness of putative last block (end at R)
        arg_log <- block_length[1:R] - block_length[R+1]
        #arg_log[arg_log <= 0] <- Inf

        nn_cum_vec <- cumsum(nn_vec[R:1])
        nn_cum_vec <- nn_cum_vec[R:1]
        
        fit_vec <- nn_cum_vec * (log(nn_cum_vec) - log(arg_log))
        fit_vec <- fit_vec - prior
        
        supp <- c(0, best) + fit_vec
        
        best <- c(best, max(supp))
        last <- c(last, which.max(supp))
         
}    
    
    #---------------------------------------------
    # Now find changepoints by iteratively peeling off the last block
    #---------------------------------------------
    index <- last[N]

    change_points <- NULL

    while(index > 1){
        change_points <- c(index, change_points)
        index <- last[index - 1]
    }
    
#     change_points <- c(change_points, blocks[1])
    
    return(c(begin, blocks[change_points], end))
}
```

```{r}
test <- c(rnorm(1000, 0, 1), rgamma(1000, 9, 2))
test <- test[(test > -5) & (test < 10)]
test <- sort(test)
```

```{r}
cp_h <- bayesian_blocks(test, prior=7.61)
cp_h
```

```{r}
options(repr.plot.width=18, repr.plot.height=8) 
par(mfrow=c(1,2))

h_t  <- hist(test, breaks=200)

h_bb <- hist(test, breaks=cp_h )

```

Prior studies

The needed ncp_prior–p0 relationship is easily found by
noting that the rates of correct and incorrect responses to
fluctuations in simulated pure noise can be controlled by
adjusting the value of ncp_prior. The procedure is: generate a
synthetic pure noise time series; apply the algorithm for a range
of ncp_prior; and select the smallest value that yields false
detection frequency equal or less than the desired rate, such
as 0.05. The values of ncp_prior determined in this way are
averaged over a large number of realizations of the random data.
The result depends on only the number of data points and the
adopted value of p0:

```{r}

```

```{r}

```

Import data

```{r}
data   <- read.table("./Data/B19036_AmCsCo_20180316.dat", skip=2)
counts <- data[,1]
bins   <- c(0:8191)

# Preliminary computation: start data at the first non-empty bin

idx    <- which(counts >0)[1]
counts <- counts[idx:length(counts)]
bins   <- bins[idx:length(bins)]
```

```{r}
options(repr.plot.width=6, repr.plot.height=4)  #to set graph size

plot(bins, counts, type='s')#, xlim=c(6520,6545))

```

Prior

```{r}
ncp_prior <- c(1, 1.3, 2, 2.5, 3.2, 4)
#ncp_prior <- 7.61

```

Algorithm

```{r}
#create histogram

rebin_bb <- function(bins, counts, change_points){
    
    rebin <- NULL
    y <- 0
    dn <- bins[2]-bins[1]
    n <- 1
    
    N <- length(bins)
    start <- bins[1]
    stop <- bins[N]

    mids <- c(0.5*(bins[2:N]+bins[1:N-1]), stop)

    for (i in 1:(length(counts))){
        y <- y + (dn*counts[i])
        ifelse( mids[i] %in% change_points,
                {y <- y/n
                 rebin <- c(rebin, y)
                 y <- 0
                 n <- 1},
                 n <- n+1
        )
    }
    return(rebin)
}
```

```{r}
#create histogram

rebin_bb <- function(bins, counts, change_points){
    
    rebin <- NULL
    y <- 0
    dn <- bins[2]-bins[1]
    
    for (i in 1:(length(change_points)-1)){
        ifelse( ( bins < change_points[i+1] ) & ( bins > change_points[i] ),
               { y <- sum(counts[bins < change_points[i+1] & bins > change_points[i]])
                 y <- y/(dn*length(counts[bins < change_points[i+1] & bins > change_points[i]]))
                 rebin <- c(rebin, y)},
               next
            )
    }
    return(c(0,rebin,0))
}
```

```{r}
cp <- bayesian_blocks(xs = bins, prior=1, nn_vec = counts)
rebin <- rebin_bb(bins = bins, counts = counts, change_points = cp)
```

```{r}
plot(bins, counts, col= 'grey', 
     type='s', lwd=0.1, log='y')
lines(cp, rebin, type='s', col='red')
#abline(v=cp, col='red')


```

```{r}

```

```{r}

```

```{r}

```

```{r}
par(mfrow=c(3,2), mar=c(3.5,3.5,0.5,0.5), oma=c(0.1,0.1,0.1,0.5), mgp=c(2.0,0.8,0))

for(i in 1:length(ncp_prior)) {
    cp <- bayesian_blocks(blocks = block_length, data = nn_vec, prior = ncp_prior[i])
    h  <- rebin_bb(data = data[,1], change_points = cp)
    plot(block_length[length(block_length):2], data[,1], col= 'grey', 
         type='s', log='y', lwd=0.1, main=ncp_prior[i])
    lines(cp, h, col='red', type='s', lwd=2)
}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}
# ---------------------------------------------
# Start with first data cell; add one cell at
# each iteration
# ---------------------------------------------
best <- NULL 
last <- NULL
supp <- NULL

for (R in 1:8193){
# Compute fit_vec : fitness of putative last block (end at R)
    arg_log <- block_length[1:R] - block_length[R+1]
    arg_log[arg_log <= 0] <- Inf
    
    nn_cum_vec <- cumsum(nn_vec[R:1])
    nn_cum_vec <- nn_cum_vec[R:1]
    
    fit_vec <- nn_cum_vec * (log(nn_cum_vec) - log(arg_log))
    
    supp <- c(0, best) + fit_vec - ncp_prior

    best <- c(best, max(supp))
    last <- c(last, which.max(supp))
}    
```

```{r}
last
```

```{r}
# #---------------------------------------------
# # Now find changepoints by iteratively peeling
# off the last block
# #---------------------------------------------
index <- last[length(nn_vec)]
change_points <- NULL

while (index > 1){
    change_points <- c(index, change_points)
    index <- last[index - 1]
}
change_points <- c(change_points, 8191)
```

```{r}
plot(block_length[length(block_length):2], data[,1], col= 'red', type='s', log='y')
lines(change_points, c(rebin, rebin[length(rebin)]), col='green', type='s')

#plot(change_points, c(rebin, rebin[length(rebin)]), col='green', type='s')

```

```{r}

```

```{r}
bayesian_blocks <- function(blocks, data, prior=7.61){
    
    #data <- sort(data)
    #N <- length(data)
    
    # ---------------------------------------------
    # Start with first data cell; add one cell at
    # each iteration
    # ---------------------------------------------
    
    best <- NULL 
    last <- NULL
    supp <- NULL

    for (R in 1:length(blocks)){
    # Compute fit_vec : fitness of putative last block (end at R)
        arg_log <- blocks[1:R] - blocks[R+1]
        arg_log[arg_log <= 0] <- Inf
        
        nn_cum_vec <- cumsum(data[R:1])
        nn_cum_vec <- nn_cum_vec[R:1]
        
        fit_vec <- nn_cum_vec * (log(nn_cum_vec) - log(arg_log))
       
        supp <- c(0, best) + fit_vec - prior
        
        best <- c(best, max(supp))
        last <- c(last, which.max(supp))
         
}    

    # #---------------------------------------------
    # # Now find changepoints by iteratively peeling
    # off the last block
    # #---------------------------------------------
    index <- last[length(data)]
    change_points <- NULL

    while(index > 1){
        change_points <- c(index, change_points)
        index <- last[index - 1]
    }
    
    change_points <- c(change_points, blocks[1])
    
    return(change_points)
}
```

```{r}
#create histogram

rebin_bb <- function(data, change_points){
    rebin <- NULL
    y <- NULL
    n <- 1

    for (i in 1:length(data)){
        y <- y + data[i]
        ifelse( i %in% change_points,
               {y <- y/n
                rebin <- c(rebin, y)
                y <- 0
                n <- 1},
                n <- n+1
        )
    }
    return(c(rebin, rebin[length(rebin)]))
}
```
